New cardinality estimation algorithms for HyperLogLog sketches

### 2016-06-15 paper/paper.tex
reformatted some equations

Using the identity $2\symSmallCorrectionFunc(\symX) + \symLargeCorrectionFunc(\symX) = -\symPowerSeriesFunc(\log_2(-\log(\symX)))/\left(\log(2)\log(\symX)\right)$, we get for the linear counting case with $\symRegRange=0$
\begin{equation}
\symPoissonRateEstimate 
= 
\frac{\symNumReg}{\log(2)\left(
2\symSmallCorrectionFunc(\symCount_0/\symNumReg)
+
\symLargeCorrectionFunc(\symCount_0/\symNumReg)
\right)}
=
\frac{\symNumReg\log\!\left(\frac{\symNumReg}{\symCount_0}\right)}{\symPowerSeriesFunc\!\left(\log_2\!\left(\log\!\left(\frac{\symNumReg}{\symCount_0}\right)\right)\right)}
\approx
\symNumReg\log\!\left(\symNumReg/\symCount_0\right),
\end{equation}
which is almost identical to the linear counting estimator used for the small range correction \eqref{equ:small_range_estimate}. Here we used the fact that $\symPowerSeriesFunc$ can be well approximated by 1.
\subsection{Corrected raw estimation algorithm}
The corrected raw estimator \eqref{equ:correctedestimator} leads to a new cardinality estimation algorithm for HyperLogLog sketches. \cref{alg:corrected_raw_estimation} demonstrates the numerical evaluation of the estimator. The values of functions $\symSmallCorrectionFunc$ and $\symLargeCorrectionFunc$ can be either calculated on-demand or pre-calculated. The possible value range for $\symCount_0$ and $\symCount_{\symRegRange+1}$ is $[0, \symNumReg]$. Therefore, if performance matters, it is possible to precalculate the function values for all possible arguments and keep them in lookup tables of size $\symNumReg+1$.

>>>>

Using the identity $2\symSmallCorrectionFunc(\symX) + \symLargeCorrectionFunc(\symX) = \symPowerSeriesFunc(\log_2(\log(1/\symX)))/\left(\log(2)\log(1/\symX)\right)$, we get for the linear counting case with $\symRegRange=0$
\begin{equation}
\symPoissonRateEstimate 
= 
\frac{\symNumReg}{\log(2)\left(
2\symSmallCorrectionFunc(\symCount_0/\symNumReg)
+
\symLargeCorrectionFunc(\symCount_0/\symNumReg)
\right)}
=
\frac{\symNumReg\log\!\left(\symNumReg/\symCount_0\right)}{\symPowerSeriesFunc\!\left(\log_2\!\left(\log\!\left(\symNumReg/\symCount_0\right)\right)\right)}
\approx
\symNumReg\log\!\left(\symNumReg/\symCount_0\right),
\end{equation}
which is almost identical to the linear counting estimator used for the small range correction \eqref{equ:small_range_estimate}. Here we used the fact that $\symPowerSeriesFunc$ can be well approximated by 1.
\subsection{Corrected raw estimation algorithm}
The corrected raw estimator \eqref{equ:correctedestimator} leads to a new cardinality estimation algorithm for HyperLogLog sketches. \cref{alg:corrected_raw_estimation} demonstrates the numerical evaluation of the estimator. The values of functions $\symSmallCorrectionFunc$ and $\symLargeCorrectionFunc$ can be either calculated on-demand or pre-calculated. The possible value range for $\symCount_0$ and $\symCount_{\symRegRange+1}$ is $[0, \symNumReg]$. Therefore, if performance matters, it is possible to precalculate the function values for all possible arguments and keep them in lookup tables of size $\symNumReg+1$.
